### 2018-01-27 supplementary_info/Supplementary_figures.Rmd
incorporating final comments

---
output:
  pdf_document:
    fig_caption: yes
---

## Supplementary Figures
\setcounter{figure}{0}
\setcounter{table}{0}
\makeatletter
\renewcommand{\thefigure}{S\@arabic\c@figure}
\renewcommand{\thetable}{S\@arabic\c@table}
\makeatother

####Agalma analysis

![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](Supplementary_Figure1.png)

\pagebreak

####SOWH analysis

![Constrained topologies specified in SOWH testing. Test statistic and p-value for each tree estimated under constraint are given.](Supplementary_Figure2.png)

\pagebreak

#####Stochastic Character maps

![Stochastic character map of presence of tentilla with *Physalia* included as not bearing tentilla.](Supplementary_Figure3.pdf)

![Stochastic character map of presence of tentilla with *Physalia* included as bearing tentilla.](Supplementary_Figure4.pdf)

![Stochastic character map of presence of nectosome.](Supplementary_Figure5.pdf)

![Stochastic character map of presence of a descending mantle canal in the nectophores. Cystonects and Athorybia were excluded as they do not have a nectosome.](Supplementary_Figure6.pdf)

![Stochastic character map for the evolution of the position of the nectosome. Cystonects were excluded as they do not have a nectosome.](Supplementary_Figure7.pdf)

![Brownian Motion character map of median depth of species observed with an MBARI ROV](Supplementary_Figure8.pdf)

![Brownian Motion character map of median depth of species including blue water diving observations.](Supplementary_Figure9.pdf)